A weakly structured stem for human origins in Africa

### 2023-02-26 paper.tex
add references

We also find that Stem 2 continued to contribute to western Africans during the
Last Glacial Period, indicative that this gene flow likely occurred in
western/central Africa (Table~\ref{tab:migration-rates}). Such an
interpretation is reinforced by the differential migration rates between
regions, i.e., the gene flow from Stem 2 to western Africans is estimated to be
five times that of the rate to eastern Africans during this period. We
performed a variety of validation tests to explore sensitivity of our
assumptions, including relaxing fixed parameters (SI
Section~\ref{sec:validation_strategy}). Most validation tests resulted in
parameters similar to the models discussed above. However, one exception was
the inferred Out-of-Africa and eastern/western African divergences which were
10--15ka younger than our fixed parameters. These younger dates are clearly at 
odds with the accepted timing of Out-of-Africa expansion at approximately $\sim>50$ka, 
as based on archaeological, climatic and paleo-morphological information [cite]. 
We interpret the free estimate for eastern African vs. European divergence as reflecting 
our inclusion of only a single OOA population in the model, the lack of a nearby 
source for back-to-Africa gene flow, and other regionally complex parameters. 
However, older pan-African features of our inferred models are weakly affected by 
this analysis choice (SI Section~\ref{sec:models}).

>>>>

We also find that Stem 2 continued to contribute to western Africans during the
Last Glacial Period, indicative that this gene flow likely occurred in
western/central Africa (Table~\ref{tab:migration-rates}). Such an
interpretation is reinforced by the differential migration rates between
regions, i.e., the gene flow from Stem 2 to western Africans is estimated to be
five times that of the rate to eastern Africans during this period. We
performed a variety of validation tests to explore sensitivity of our
assumptions, including relaxing fixed parameters (SI
Section~\ref{sec:validation_strategy}). Most validation tests resulted in
parameters similar to the models discussed above. However, one exception was
the inferred Out-of-Africa and eastern/western African divergences which were
10--15ka younger than our fixed parameters. These younger dates are clearly at 
odds with the accepted timing of Out-of-Africa expansion at approximately $\sim>50$ka, 
as based on archaeological, climatic and paleo-morphological information
\citep{Fu2014-am,Groucutt2015-do,Prufer2021-ao,Beyer2021-ym}. 
We interpret the free estimate for eastern African vs. European divergence as reflecting 
our inclusion of only a single OOA population in the model, the lack of a nearby 
source for back-to-Africa gene flow, and other regionally complex parameters. 
However, older pan-African features of our inferred models are weakly affected by 
this analysis choice (SI Section~\ref{sec:models}).
